---
title: "Deliverable 3"
output: html_notebook
---

Loading in the data
```{r}
rm(list = ls())
library(rio)

location ='https://github.com/reederan/542-finaldeliverable-/raw/main/'
file ='seattleTop.csv'
delivdata = paste0(location,file)

delivdata2 = read.csv(delivdata)

```

Checking the dimensions of the data frame to create a matrix
```{r}
dim(delivdata2)

names(delivdata2)

row.names(delivdata2)=delivdata2$name
delivdata2$name=NULL

dim(delivdata2)
```

Converting into matrix
```{r}
delivdata_mx=as.matrix(delivdata2) 
```

Creating a network
```{r}

library(igraph)

EliteNet=graph.adjacency(delivdata_mx,mode="undirected",weighted=NULL)

```

Identifying how many vertices the network has
```{r}
vcount(EliteNet)
```

Identifying how many edges the network has
```{r}
ecount(EliteNet)
```

looking at the network
```{r}
plot.igraph(EliteNet,
            vertex.color = 'yellow',
            edge.color='lightblue')
```

Seeing if network is connected
```{r}
is_connected(EliteNet)
```

looking at the giant component (do i need to do this part since the network is connected?????)
```{r}
EliteNet_giant=CINNA::giant_component_extract(EliteNet,directed = F)[[1]]

plot.igraph(EliteNet_giant)

summary(EliteNet_giant)
```

looking at the graph density
```{r}
graph.density(EliteNet_giant)
```

```{r}
diameter(EliteNet_giant)
```

```{r}
farthest_vertices(EliteNet_giant)
```

```{r}
average.path.length(EliteNet_giant)
```

checking transitivity
```{r}
transitivity(EliteNet_giant)
```

```{r}
assortativity_degree(EliteNet_giant)
```

```{r}
rounding=3
degr=round(degree(EliteNet_giant,normalized=T),rounding)
close=round(closeness(EliteNet_giant,normalized=T),rounding)
betw=round(betweenness(EliteNet_giant,normalized=T),rounding)
eigv=round(eigen_centrality (EliteNet_giant,scale=F)$vector,rounding)

DFCentrality=as.data.frame(cbind(close,betw,eigv),stringsAsFactors = F)
names(DFCentrality)=c('Closeness','Betweenness','Eigenvector')
```

```{r}
DFCentrality$person=row.names(DFCentrality)
row.names(DFCentrality)=NULL
head(DFCentrality)
```

```{r}
library(ggplot2)
ggplot(DFCentrality, aes(x=Betweenness, y=Closeness)) + 
    theme_classic() +
    geom_text(aes(label=person,color=Eigenvector)) +
    scale_colour_gradient(low = "orange", high = "black")
```

```{r}
vertex_connectivity(EliteNet_giant)
```

```{r}
(TheCut=articulation_points(EliteNet_giant))
```

```{r}
(TheBridge=bridges(EliteNet_giant))
```

```{r}
allSizes=rep(10,vcount(EliteNet_giant))
allSizes[TheCut]=40
V(EliteNet_giant)$size=allSizes 
plot.igraph(EliteNet_giant,layout = layout.fruchterman.reingold,
             vertex.color = 'yellow',
             edge.color='lightblue',vertex.shape='sphere')
```

```{r}
E(EliteNet_giant)$color=ifelse(E(EliteNet_giant)==TheBridge,
                               'red','grey80')
                               
plot.igraph(EliteNet_giant,layout = layout.fruchterman.reingold,
             vertex.color = 'yellow',
             edge.color=E(EliteNet_giant)$color,vertex.shape='sphere')
```

```{r}
write_graph(EliteNet, "EliteNetR.graphml", "graphml")
```

